\chapter{Previous works}
The internet is brimming with applications designed to enhance board games, and so are the articles and projects expanding upon this field. Similarly to this thesis, a mobile library was proposed by Takács in his thesis from 2025~\cite{takacs}, where an Android mobile library was suggested for so called \textbf{smart assistants}, applications on mobile devices that help with playing board games, similar to this thesis. The overall intrusiveness, or rather, how much these applications affect the players can differ, as each game and its players may have different needs.

If these applications were grouped into categories, the first category would be applications serving as a digital copy of the game rules.

Another such category would be \textbf{management applications} that help players organize their data related to the game. These applications may not react to the game's state but rather provide users with a way to track their game. A great example of such an application is \textit{D\&D Beyond}, a mobile application for \textit{Dungeons \& Dragons}. This application contains information about the player's character and essentially replaces the character sheet, a paper form of this data.

The next category is \textbf{move suggestion} applications. These applications are usually aimed at helping newer players learn the game faster or giving players an advantage over others. An example of such an application would be \textit{CubeX}, a Rubik's cube assistant that lets players scan or manually fill in the required side's of the cube, and the application will suggest moves. Similarly, Grossman, in his thesis from 2018~\cite{grossman}, implemented a move suggestion application for the game \textit{Krycí Jména}, giving users an edge over other players. Another thesis by Široký~\cite{siroky} focused on \textit{Checkers, Gomoku, and Nine men’s morris}, while utilizing classical computer vision instead of CNNs.

The last category of applications is game \textbf{enhancers}. These applications usually expand the base game by providing additional functionalities or are mandatory to play the game. An example of such a game is \textit{X-COM: The Board Game}, where the application is mandatory and forces time limited decisions onto the players.