% This file should be replaced with your file with thesis content.
%=========================================================================
% Authors: Michal Bidlo, Bohuslav Křena, Jaroslav Dytrych, Petr Veigend and Adam Herout 2019

% For compilation piecewise (see projekt.tex), it is necessary to uncomment it and change
% \documentclass[../projekt.tex]{subfiles}
% \begin{document}
\chapter{Experiments and evaluation}
This chapter shows experiments performed on gathered image data to train the object detection model for the Bang! card. Furthermore, it evaluates the proposed Dataset Creator and CVLiBG.

%=========================================================================

%=========================================================================

\section{Experiment -- training models for the Bang! card game}
\label{bangTrainingExperiments}
This section compares models trained on different datasets and their export options to convert models into the ONNX format. The dataset table \ref{datasetTable} shows different datasets used during testing. All training was done in Dataset Creator with available models from Ultralytics, with the patience value set to 50, meaning that if the model did not improve after 50 epochs, the training would be suspended. The synthetic data used in the experiments were generated using the Dataset Creator. All experiments were performed on consumer-grade hardware. The hardware configurations can be seen in table \ref{hardwareConfiguration}.

\begin{table}[hbt]
\caption{Used hardware for experiments.}
\centering
\label{hardwareConfiguration}
\begin{tabular}{|cc|}
\hline
\multicolumn{2}{|c|}{Model training}                            \\ \hline
\multicolumn{1}{|c|}{CPU}        & AMD Ryzen 4600H               \\ \hline
\multicolumn{1}{|c|}{GPU}        & NVIDIA RTX 2060 Mobile        \\ \hline
\multicolumn{1}{|c|}{GPU Memory} & 6GB                           \\ \hline
\multicolumn{1}{|c|}{Memory}     & 16GB                          \\ \hline
\multicolumn{2}{|c|}{Mobile benchmarks}                          \\ \hline
\multicolumn{1}{|c|}{CPU}        & Mediatek Dimensity 7300 Ultra \\ \hline
\multicolumn{1}{|c|}{GPU}        & Mali-G615 MC2                 \\ \hline
\multicolumn{1}{|c|}{Memory}     & 12GB                          \\ \hline
\end{tabular}
\label{datasetTable}
\caption{Table of used dataset for the experiments. The \textbf{xN} represents a \textbf{N} synthetic data generated per image, \textbf{+ org} refers to unmodified, original, data was used as well.}
\centering
\begin{tabular}{|c|c|c|c|c|c|}
\hline
Dataset       & Shortcut & Train & Validation & Synthetic data & Resized \\ \hline
Alpha         & A        & 539   & 197        & Org        & No\\ \hline
Beta          & B        & 2156  & 788        & x3 + Org   & No\\ \hline
Gamma         & G        & 3773  & 1379       & x6 + Org   & No\\ \hline
Gamma Resized & GR       & 3773  & 1379       & x6 + Org   & Yes\\ \hline
Delta         & D        & 3773  & 1379       & x6         & Yes\\ \hline
Alpha Simplified & AS       & 90    & 639     & Org        & No\\ \hline
Gamma Simplified & GS       & 570   & 639     & x6         & No\\ \hline
\end{tabular}
\end{table}


\subsection{Understanding the results}
The next subsections will show result tables with data. To correctly understand what each column means, a quick review was prepared \cite{ultralyticsResults}:

\begin{itemize}
    \item {\textbf{time} cumulative training time elapsed up to the given epoch. This value is represented in seconds.}
    \item {\textbf{epoch} is one complete pass of the model over the entire training dataset. Each epoch ends with an evaluation on the validation part of the dataset.}
    \item {\textbf{t/} columns representing evaluation metrics computed on the training dataset, indicating how well the model fits the data it was trained on.}
    \item {\textbf{v/} columns representing evaluation metrics computed on the validation dataset, meaning it provides an estimate of the model's ability to recognize unseen data.}
    \item {\textbf{box\_loss} measures the error in predicted bounding boxes. Lower values indicate more accurate object localization.}
    \item {\textbf{cls\_loss} measures the error in class predictions. Lower values indicate better classification performance.}
    \item {\textbf{dfl\_loss} (Distribution Focal Loss) measures the accuracy of bounding box edge regression by modeling bounding box coordinates as probability distributions. Lower values indicate more precise localization.}
    \item {\textbf{precision} represents the proportion of correctly predicted objects. Higher values indicate the ability to recognize correct detections.}
    \item {\textbf{recall} represents the proportion of ground truth objects detected in the image. High recall indicates that the model successfully detects most objects.}
    \item {\textbf{mAP50} is a mean Average Precision at IoU = 0.50. This evaluates detection performance with a less strict overlap threshold. Higher values indicate better detection accuracy.}
    \item {\textbf{mAP50-95} is a mean Average Precision averaged over IoU thresholds from 0.50 to 0.95, which provides a stricter evaluation of detection performance.}
\end{itemize}

\subsection{Comparing Open Neural Network Exchange with OpenCV detectors}
CVLiBG internally uses OpenCV for image manipulation. However, it also implements an additional module with ONNX based detectors. ONNX was chosen as an OpenCV alternative due faster inference latency. This experiment show measured differences between the two inference runtimes.

\begin{table}[hbt]
\caption{Table showing difference between Open Neural Network Runtime in comparison with OpenCV's DNN runtime.}
\centering
\begin{tabular}{|ccccccc|}
\hline
\multicolumn{1}{|c|}{} &
  \multicolumn{1}{c|}{Total} &
  \multicolumn{1}{c|}{Preproc} &
  \multicolumn{1}{c|}{Conversion} &
  \multicolumn{1}{c|}{Inference} &
  \multicolumn{1}{c|}{Thresholding} &
  NMS \\ \hline
\multicolumn{7}{|c|}{ONNX} \\ \hline
\multicolumn{1}{|c|}{average} &
  \multicolumn{1}{c|}{134.3490} &
  \multicolumn{1}{c|}{0.5555} &
  \multicolumn{1}{c|}{10.9532} &
  \multicolumn{1}{c|}{119.9382} &
  \multicolumn{1}{c|}{2.6333} &
  0.0124 \\ \hline
\multicolumn{1}{|c|}{median} &
  \multicolumn{1}{c|}{129.3519} &
  \multicolumn{1}{c|}{0.5311} &
  \multicolumn{1}{c|}{10.5143} &
  \multicolumn{1}{c|}{115.0735} &
  \multicolumn{1}{c|}{2.5279} &
  0.0111 \\ \hline
\multicolumn{7}{|c|}{OpenCV} \\ \hline
\multicolumn{1}{|c|}{average} &
  \multicolumn{1}{c|}{245.5628} &
  \multicolumn{1}{c|}{0.5349} &
  \multicolumn{1}{c|}{3.6952} &
  \multicolumn{1}{c|}{236.2461} &
  \multicolumn{1}{c|}{3.4799} &
  0.4261 \\ \hline
\multicolumn{1}{|c|}{median} &
  \multicolumn{1}{c|}{238.1550} &
  \multicolumn{1}{c|}{0.5226} &
  \multicolumn{1}{c|}{3.1983} &
  \multicolumn{1}{c|}{228.9654} &
  \multicolumn{1}{c|}{3.1094} &
  0.3894 \\ \hline
\end{tabular}
\end{table}

OpenCV's implementation benefits from its image \textbf{conversion} being, on average, 7 milliseconds faster. However, the \textbf{inference} latency shows a difference of almost 113 milliseconds, being almost 2x slower.

\subsection{Limited dataset with and without synthetic data}
This test shows the difference between a dataset with no synthetic data and a dataset with \textbf{x6} synthetic data on very limited image data, mostly 2 images per class, for a total of 45 classes. Both models were trained on YOLO version 8 Nano.

\begin{table}[hbt]
\caption{Alpha simplified dataset in comparison with Gamma simplified dataset. Except time and epochs, higher values are preferred.}
\centering
\begin{tabular}{|c|c|c|c|c|c|c|}
\hline
Dataset    & epochs   & time  & precision & recall & mAP50 & mAP50-95 \\ \hline
AS         & 136      & 17.94 & 0.54506      & 0.46305   & 0.48355  & 0.47047     \\ \hline
GS         & 29       & 8.67  & 0.56125      & 0.52848   & 0.555    & 0.54066     \\ \hline
$\Delta$\% & 129.6970 & 69.65 & 2.9268       & 13.1978   & 13.7596  & 13.8835     \\ \hline
\end{tabular}
\end{table}

\begin{table}[hbt]
\caption{Alpha Simplified in comparison with Gamma Simplified. Lower values are preferred.}
\centering
\begin{tabular}{|c|c|c|c|c|c|c|}
\hline
Dataset & t/box\_loss & t/cls\_loss & t/dfl\_loss & v/box\_loss & v/cls\_loss & v/dfl\_loss \\ \hline
AS         & 0.23063 & 0.91337 & 0.86579 & 0.23749 & 2.03006 & 0.84417 \\ \hline
GS         & 0.25085 & 0.94998 & 0.89367 & 0.23902 & 1.71815 & 0.85772 \\ \hline
$\Delta$\% & 8.3991  & 3.9295  & 3.1692  & 0.6422  & 16.6431 & 1.5923  \\ \hline
\end{tabular}
\end{table}



Models were set to train for 1000 epochs; however, as the models did not improve, the training was suspended automatically, and the best results were used for evaluation. The \textbf{GS} dataset with synthetic data was trained for 70\% less time (in this case, 9 minutes), while the mean average precision (mAP) is 13.8\% higher. This experiment shows that synthetic data can help with models trained on very limited datasets while achieving higher accuracy and lower training time.

\subsection{Comparing datasets}
The next experiment shows the difference between datasets. As datasets have different sizes, it was not possible to compare them based on the trained epochs. Therefore, a time was used to compare datasets. All datasets were trained on YOLO version 26.

\begin{table}[hbt]
\centering
\caption{Comparison of datasets with similar time. Higher values are preferred.}
\label{expDatasetComparison1}
\begin{tabular}{|l|l|l|l|l|l|l|}
\hline
Dataset & epoch & time & precision & recall & mAP50 & mAP50-95 \\ \hline
A  & 146 & 2867.36 & 0.91806 & 0.94451 & 0.97659 & 0.97063 \\ \hline
B  & 40  & 2850.03 & 0.93823 & 0.91781 & 0.97539 & 0.96963 \\ \hline
D  & 37  & 2835.91 & 0.91314 & 0.92936 & 0.97559 & 0.97147 \\ \hline
G  & 23  & 2857.56 & 0.96487 & 0.95248 & 0.98537 & 0.98476 \\ \hline
GR & 33  & 2872.23 & 0.95742 & 0.93712 & 0.97581 & 0.97474 \\ \hline
\end{tabular}
\end{table}

\begin{table}[hbt]
\centering
\label{expDatasetComparison2}
\caption{Comparison of datasets with similar time. Lower values are preferred.}
\begin{tabular}{|l|l|l|l|l|l|l|}
\hline
Dataset & t/box\_loss & t/cls\_loss & t/dfl\_loss & v/box\_loss & v/cls\_loss & v/dfl\_loss \\ \hline
A         & 0.19101 & 0.79243 & 0.00298 & 0.17482 & 0.23966 & 0.00473 \\ \hline
B         & 0.15251 & 1.24147 & 0.00316 & 0.15994 & 0.25554 & 0.00442 \\ \hline
D         & 0.1507  & 0.69188 & 0.00314 & 0.15564 & 0.25739 & 0.00315 \\ \hline
G         & 0.21079 & 0.76429 & 0.0031  & 0.1523  & 0.16473 & 0.00395 \\ \hline
GR        & 0.14538 & 0.54868 & 0.00302 & 0.15805 & 0.21347 & 0.00316 \\ \hline
\end{tabular}
\end{table}

The first table \ref{expDatasetComparison1} shows that the \textbf{Gamma} and \textbf{Gamma Resized} datasets perform better than other datasets, mostly due to more variance in synthetic data. While the non resized dataset shows slightly higher precision.

The second table \ref{expDatasetComparison2}, on the other hand, favors the \textbf{Gamma Resized} dataset in training \texttt{box\_loss} a \texttt{cls\_loss}. These differences are 36\% and 31\%; however, these metrics show the training data. More relevant data are validation losses and metrics from the \ref{expDatasetComparison1}, which favor the \textbf{Gamma} dataset. These results suggest that the \textbf{Gamma Resized} is performing better on training data; however, it is worse in generalization, specifically in the validation part of the dataset.

The \textbf{Gamma} dataset is performing better than other datasets and is better at generalization, mostly due to the greater amount of synthetic data, which allows the model to learn the features of the objects. And therefore, it was chosen for further testing and final deployment.

\subsection{Comparing accuracy of YOLO versions}
Different versions of the same model family behave differently; as their architectures change, so do their performance and accuracy. Therefore, three different versions of YOLO models were tested. All models are using the pretrained \textit{Nano} version from Ultralytics.

\begin{table}[hbt]
\label{expYoloVersions1}
\caption{Comparison of YOLO versions. Higher values are preferred.}
\centering
\begin{tabular}{|c|c|c|c|c|c|c|}
\hline
Version                  & epoch    & time     & precision & recall  & mAP50   & mAP50-95 \\ \hline
8                               & 89       & 11959.9  & 0.97643   & 0.97336 & 0.98958 & 0.9804   \\ \hline
11                              & 35       & 5394.61  & 0.95968   & 0.97174 & 0.97876 & 0.96871  \\ \hline
26                              & 57       & 9500.06  & 0.97766   & 0.96016 & 0.98154 & 0.96982  \\ \hline
8 to 26 $\Delta$\% & 43.83562 & 22.92493 & 0.12589   & 1.36539 & 0.81578 & 1.08501  \\ \hline
\end{tabular}
\end{table}

\begin{table}[hbt]
\label{expYoloVersions2}
\caption{Comparison of YOLO versions. Lower values are preferred.}
\centering
\begin{tabular}{|c|c|c|c|c|c|c|}
\hline
Model versions & t/box\_loss & t/cls\_loss & v/box\_loss & v/cls\_loss \\ \hline
8              & 0.23209         & 0.32786         & 0.1541        & 0.14738       \\ \hline
11             & 0.28223         & 0.46873         & 0.17432       & 0.19751       \\ \hline
26             & 0.33584         & 0.38488         & 0.17799       & 0.13504       \\ \hline
$\Delta$\% difference between 8 and 26      & 36.53619        & 16.00022        & 14.38767      & 8.73876       \\ \hline
\end{tabular}
\end{table}

Tables \ref{expYoloVersions1} and \ref{expYoloVersions2} compare YOLO versions 8, 11, and 26. Based on the metrics from the table \ref{expDatasetComparison1}, the best performing models are YOLO 8 and YOLO 26, with YOLO 8 being slower to train but \textasciitilde 1\% more accurate in \texttt{mAP50} and \texttt{mAP50-95} metrics. Table \ref{expYoloVersions2} shows similar results where YOLO 8 is outperforming other YOLO versions, except in validation class loss.

\subsection{Comparing inference latency on different export architectures}
Object detection models in this project are used on mobile devices with limited resources. Therefore, a combination of accuracy and latency needs to be considered. The \ref{expYoloVersionsSpeed} table shows the runtime performance of each tested model with different optimizations. Trained models can be exported for different architectures. While it is generally accepted that GPU and NPU are preferred for machine learning and neural networks, more budget oriented devices might not utilize these features. A great example would be mobile devices used for this testing; performing object detection tasks using the Neural Networks API (NNAPI) was 4x slower than running these models on a device's CPU. 

These values might differ based on the device's hardware capabilities. However, based on these results, it is recommended to include multiple models and allow users to either switch between them or create an automatic solution. Different model architectures and optimizations are generated by the Dataset Creator on export.

Further, this table shows a significant difference in inference latency between versions, with an average difference of 20 milliseconds between the YOLO 8 and YOLO 26 versions. 

\begin{table}[hbt]
\caption{Comparison of inference latency with different export and Detector architectures.}
\label{expYoloVersionsSpeed}
\centering
\begin{tabular}{|cccc|}
\hline
\multicolumn{1}{|c|}{}        & \multicolumn{1}{c|}{CPU}     & \multicolumn{1}{c|}{GPU}     & NPU     \\ \hline
\multicolumn{4}{|c|}{YOLO 8}                                                                          \\ \hline
\multicolumn{1}{|c|}{average} & \multicolumn{1}{c|}{134.356} & \multicolumn{1}{c|}{171.141} & 578.837 \\ \hline
\multicolumn{1}{|c|}{median}  & \multicolumn{1}{c|}{130.475} & \multicolumn{1}{c|}{167.663} & 575.477 \\ \hline
\multicolumn{4}{|c|}{YOLO 11}                                                                         \\ \hline
\multicolumn{1}{|c|}{average} & \multicolumn{1}{c|}{122.011} & \multicolumn{1}{c|}{124.901} & 535.586 \\ \hline
\multicolumn{1}{|c|}{median}  & \multicolumn{1}{c|}{118.857} & \multicolumn{1}{c|}{121.730} & 534.213 \\ \hline
\multicolumn{4}{|c|}{YOLO 26}                                                                         \\ \hline
\multicolumn{1}{|c|}{average} & \multicolumn{1}{c|}{110.331} & \multicolumn{1}{c|}{112.087} & 498.623 \\ \hline
\multicolumn{1}{|c|}{median}  & \multicolumn{1}{c|}{106.385} & \multicolumn{1}{c|}{109.686} & 495.079 \\ \hline
\end{tabular}
\end{table}

\subsection{Overall conclusion}
Based on the measured results, the most accurate model is \textbf{YOLO 8} trained on the \textbf{Gamma} dataset. However, the best performing model, when considering both accuracy and latency, is \textbf{YOLO 26} trained on the \textbf{Gamma} dataset.

The best export architecture was the \textbf{YOLO 26 CPU}. This, however, is hardware dependent, and other devices with more suitable hardware configurations may perform better on NNAPI enabled detectors.

For the precision detection mode, where accuracy is preferred, a model trained using \textbf{YOLO 8 Medium}, instead of the \textbf{Nano} pretrained model, was used.

The \textbf{Yolo 11} was outperformed in latency and accuracy by other YOLO versions. However, it achieved the best accuracy compared to the time it took to train the model. This makes it the best for prototyping, and it is recommended to use it for Dataset Creator models for annotating data, where overall latency and accuracy might be less beneficial than the total time it took to achieve acceptable accuracy.

%=========================================================================

%=========================================================================
\section{Evaluation of the Bang! application}
\label{bangEvaluation}

The assignment of this thesis requires the developed system to be evaluated in a real-world setting during the play of the selected game. Due to time constraints, the evaluation was performed as a small-scale proof-of-concept evaluation. The goal was not to provide statistically significant results but rather to verify whether the proposed workflow can be used during gameplay and whether the developed application provides practical assistance to players.

Dataset Creator was used to create the dataset and train a custom object detection model for the Bang! card game. The trained model was then integrated into the Android application using CVLiBG. The application was tested during gameplay with the goal of observing whether it would help players by utilizing either camera based or manual search to access card information without significantly interrupting the game.

The camera based object recognition was the primary observed interaction method. Players were able to detect cards placed right in front of the camera or at angles or from a reasonable distance. Lighting conditions decreased the accuracy of the models. However, the model performed better than expected, as in some cases, due to the nature of object detection, the model was able to recognize the overall structure of the card, while the text remained hard to read for humans.

The evaluation also showed the importance of the manual search functionality. Some cards may be held by other players, placed at a distance, partially covered, or visible only for a short time. In these cases, manually searching for the card allowed the user to access the same information without interrupting the game or asking another player to show the card again.

Several limitations were observed during testing. Recognition quality depended on each card, as some cards were similar to each other, mainly weapons, which led to the wrong class being detected. The use of the application in more challenging lighting conditions led to a drop in accuracy. The application is therefore most useful when detection is combined with a manual search, rather than relying only on camera recognition. Another limitation is that the evaluation was performed on a small scale and does not provide statistically significant evidence about the improvement of the gameplay experience.

Despite these limitations, the evaluation demonstrated that the proposed tools can be used together to create a functional assistant application. The Dataset Creator supported the preparation of the object detection model for the specific game, while CVLiBG simplified its integration into the Android application. The resulting application was able to recognize selected Bang! cards, display relevant information, and provide an alternative manual access path when recognition was not possible. This confirms the practical feasibility and usability of the proposed workflow.